---
title: "Plots"
author: "Kevin Rue-Albrecht"
date: "22/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import dependencies

```{r, message=FALSE}
library(here)
library(SummarizedExperiment)
library(RColorBrewer)
library(tibble)
library(ggplot2)
library(cowplot)
library(dplyr)
library(tidyr)
```

# Import data set

```{r}
se <- readRDS(here::here('rds/se.gene_symbol.rds'))
```

# Set experimental levels

```{r}
levels.time <- c("2h", "4h", "6h", "Blank")
se$time <- factor(se$time, levels.time)
levels.infection <- c("Mock", "LT2", "D23580", "Blank")
se$infection <- factor(se$infection, levels.infection)
levels.status <- c("uninfected", "exposed", "infected", "BULK", "Blank")
se$status <- factor(se$status, levels.status)
levels.treatment <- c("Mock\nuninfected", "LT2\nexposed", "LT2\ninfected", "D23580\nexposed",  "D23580\ninfected", "Mock\nBULK", "LT2\nBULK", "D23580\nBULK",  "Blank\nBlank")
se$treatment <- factor(interaction(se$infection, se$status, sep = "\n"), levels.treatment)
```

# Define color maps

```{r}
colormap.time <- c(brewer.pal(9, "Set3")[c(2,7:8)], brewer.pal(12, "Paired")[5])
names(colormap.time) <- levels.time
colormap.infection <- brewer.pal(12, "Paired")[c(9,4,2,5)]
names(colormap.infection) <- levels.infection
colormap.status <- brewer.pal(12, "Paired")[c(9,1,7,3,5)]
names(colormap.status) <- levels.status
colormap.treatment <- RColorBrewer::brewer.pal(9, "Paired")[c(9,3,4,1,2,9,4,2,5)]
names(colormap.treatment) <- gsub("_","\n", levels.treatment)
```

# Plate layout

```{r}
plot_data <- as.data.frame(colData(se)) %>%
    separate(well, c("row", "column"), 1) %>% 
    mutate(
        row = factor(row, rev(sort(unique(row)))),
        column = factor(as.numeric(as.character(column))),
        plate = gsub("P", "Plate ", plate)
    )
```

## Time

```{r}
ggplot(plot_data) +
    geom_tile(aes(column, row, fill = time), color = 'black') +
    facet_wrap(~plate) +
    scale_fill_manual(values = colormap.time) +
    labs(
        x = NULL, y = NULL, fill = "Time"
    ) +
    theme_minimal() +
    theme(
        panel.grid.major = element_blank()
    )
```

## Infection

```{r}
ggplot(plot_data) +
    geom_tile(aes(column, row, fill = infection), color = 'black') +
    facet_wrap(~plate) +
    scale_fill_manual(values = colormap.infection) +
    labs(
        x = NULL, y = NULL, fill = "Infection"
    ) +
    theme_minimal() +
    theme(
        panel.grid.major = element_blank()
    )
```

## Status

```{r}
ggplot(plot_data) +
    geom_tile(aes(column, row, fill = status), color = 'black') +
    facet_wrap(~plate) +
    scale_fill_manual(values = colormap.status) +
    labs(
        x = NULL, y = NULL, fill = "Status"
    ) +
    theme_minimal() +
    theme(
        panel.grid.major = element_blank()
    )
```

# Subset to cells only ----

```{r}
keep <- with(
    colData(se),
    status %in% c("uninfected", "exposed", "infected"))
se.cells <- se[, keep]
```

# Number of cells per condition ----

```{r}
plot_data <- as_tibble(colData(se.cells))
```

```{r, warning=FALSE, message=FALSE}
ggplot(plot_data, aes(treatment, fill = treatment)) +
    geom_bar() +
    facet_wrap(~ time, nrow = 1) +
    labs(title = "Number of cells per condition", y = "Cells", x = "Treatment", fill = "Treatment") +
    theme_bw() +
    theme(
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank()) +
    scale_fill_manual(values = colormap.treatment)
```


# Number of genes detected over time ----

```{r}
plot_data <- tibble(
    value = colSums(assay(se.cells) > 0)
    ) %>%
    bind_cols(as.data.frame(colData(se.cells)))
```

```{r, warning=FALSE, message=FALSE}
ggplot(plot_data, aes(time, value)) +
    geom_jitter(width = 0.1, height = 0, size = 0.5) +
    facet_grid(infection ~ status) +
    geom_smooth(aes(group = interaction(status, infection), color = infection, linetype = status)) +
    labs(title = "Number of genes detected", y = "Genes", x = "Time", color = "Infection", linetype = "Status") +
    theme_bw() +
    theme(panel.grid.minor.y = element_blank()) +
    scale_color_manual(values = colormap.infection)
```

# Gene expression: box plot ----

```{r}
plot_data <- tibble(
    value = assay(se.cells)[which(rowData(se)$symbol == "IL12B"), ]
    ) %>%
    bind_cols(as.data.frame(colData(se.cells)))
```

```{r, warning=FALSE}
ggplot(plot_data, aes(time, value, fill = time)) +
    geom_boxplot() +
    facet_grid(infection ~ status) +
    scale_y_log10() +
    labs(title = "IL12B", y = "Counts", x = "Time", fill = "Time") +
    theme_bw() +
    theme(panel.grid.minor.y = element_blank()) +
    scale_fill_manual(values = colormap.time)
```

# Gene expression: violin plot ----

```{r}
plot_data <- tibble(
    value = assay(se.cells)[which(rowData(se)$symbol == "IL12B"), ]
    ) %>%
    bind_cols(as.data.frame(colData(se.cells)))
```

```{r, warning=FALSE}
ggplot(plot_data, aes(time, value, fill = time)) +
    geom_violin(draw_quantiles = 0.5, scale = "width") +
    facet_grid(infection ~ status) +
    scale_y_log10() +
    labs(title = "IL12B", y = "Counts", x = "Time", fill = "Time") +
    theme_bw() +
    theme(panel.grid.minor.y = element_blank()) +
    scale_fill_manual(values = colormap.time)
```

# Gene expression: violin plot ----

```{r}
plot_data <- tibble(
    value = assay(se.cells)[which(rowData(se)$symbol == "IL12B"), ]
    ) %>%
    bind_cols(as.data.frame(colData(se.cells)))
```

```{r, warning=FALSE, message=FALSE}
ggplot(plot_data, aes(time, value + 1)) +
    geom_jitter(width = 0.1, height = 0, size = 0.5) +
    geom_smooth(aes(group = interaction(status, infection), color = infection)) +
    scale_y_log10() +
    facet_grid(infection ~ status) +
    labs(title = "IL12B", y = "counts") +
    theme_bw() +
    theme(panel.grid.minor.y = element_blank()) +
    scale_color_manual(values = colormap.infection)
```

# Scatter plot gene 1 vs gene 2, showing samples/cells, coloured based on their annotation. ----

```{r}
plot_data <- tibble(
    IL1B = assay(se.cells)[which(rowData(se)$symbol == "IL1B"), ],
    MARCHF1 = assay(se.cells)[which(rowData(se)$symbol == "MARCHF1"), ]
    ) %>%
    bind_cols(as.data.frame(colData(se.cells)))
```

```{r, warning=FALSE}
ggplot(plot_data, aes(IL1B, MARCHF1, fill = time)) +
    geom_point(shape = 21, color = "black", stroke = 0.5, size = 2) +
    facet_grid(infection ~ status) +
    scale_y_log10() + scale_x_log10() +
    theme_bw() +
    theme(
        panel.grid.major = element_line(color = "white")) +
    scale_fill_manual(values = colormap.time)
```

